---
title: "hw 02 - Shot Charts"
author: "Justin Ng"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```


```{r Load Data}
library(dplyr)
library(tidyverse)
library(ggplot2)
options(max.print = 99999)
column_types2 <- c('character', 'character', 'integer', 'integer', 'integer', 'integer', 'character', 'character', 'character', 'integer', 'character', 'integer', 'integer')
andre_iguodala <- read.csv('../data/andre-iguodala.csv', stringsAsFactors = FALSE, colClasses = column_types2, sep = ",")
draymond_green <- read.csv('../data/draymond-green.csv', stringsAsFactors = FALSE, colClasses = column_types2, sep = ",")
kevin_durant <- read.csv('../data/kevin-durant.csv', stringsAsFactors = FALSE, colClasses = column_types2, sep = ",")
klay_thompson <- read.csv('../data/klay-thompson.csv', stringsAsFactors = FALSE, colClasses = column_types2, sep = ",")
stephen_curry <- read.csv('../data/stephen-curry.csv', stringsAsFactors = FALSE, colClasses = column_types2, sep = ",")
andre_iguodala2 <- add_column(andre_iguodala, name = 'andre_iguodala', .before = 'team_name')
draymond_green2 <- add_column(draymond_green, name = 'draymond_green', .before = 'team_name')
kevin_durant2 <- add_column(kevin_durant, name = 'kevin_durant', .before = 'team_name')
klay_thompson2 <- add_column(klay_thompson, name = 'klay_thompson', .before = 'team_name')
stephen_curry2 <- add_column(stephen_curry, name = 'stephen_curry', .before = 'team_name')
andre_iguodala2$shot_made_flag[andre_iguodala$shot_made_flag == "y"] <- "made_shot"
andre_iguodala2$shot_made_flag[andre_iguodala$shot_made_flag == "n"] <- "missed_shot"
draymond_green2$shot_made_flag[draymond_green$shot_made_flag == "y"] <- "made_shot"
draymond_green2$shot_made_flag[draymond_green$shot_made_flag == "n"] <- "missed_shot"
kevin_durant2$shot_made_flag[kevin_durant$shot_made_flag == "y"] <- "made_shot"
kevin_durant2$shot_made_flag[kevin_durant$shot_made_flag == "n"] <- "missed_shot"
klay_thompson2$shot_made_flag[klay_thompson$shot_made_flag == "y"] <- "made_shot"
klay_thompson2$shot_made_flag[klay_thompson$shot_made_flag == "n"] <- "missed_shot"
stephen_curry2$shot_made_flag[stephen_curry$shot_made_flag == "y"] <- "made_shot"
stephen_curry2$shot_made_flag[stephen_curry$shot_made_flag == "n"] <- "missed_shot"
andre_iguodala2 <- add_column(andre_iguodala2, minute = (andre_iguodala$period*12 - andre_iguodala$minutes_remaining), .after = 'minutes_remaining')
draymond_green2 <- add_column(draymond_green2, minute = (draymond_green$period*12 - draymond_green$minutes_remaining), .after = 'minutes_remaining')
kevin_durant2 <- add_column(kevin_durant2, minute = (kevin_durant$period*12 - kevin_durant$minutes_remaining), .after = 'minutes_remaining')
klay_thompson2 <- add_column(klay_thompson2, minute = (klay_thompson$period*12 - klay_thompson$minutes_remaining), .after = 'minutes_remaining')
stephen_curry2 <- add_column(stephen_curry2, minute = (stephen_curry$period*12 - stephen_curry$minutes_remaining), .after = 'minutes_remaining')

shots_data1 <- rbind(andre_iguodala2, draymond_green2, kevin_durant2, klay_thompson2, stephen_curry2)

```

```{r Question 5.1}
shots_data1 %>%
group_by(name) %>%
summarise(total_shots = length(name)) %>%
arrange(desc(total_shots))
```


```{r Question 5.2}
shots_data1 %>%
group_by(name, shot_made_flag) %>%
summarise(made_shots = length(name)) %>%
mutate(total_shots = sum(made_shots)) %>%
filter(shot_made_flag == 'made_shot') %>%
mutate(perc_made = made_shots/total_shots) %>%
select(c(1, 3:5)) %>%
arrange(desc(perc_made))

shots_data1 %>%
group_by(name, shot_made_flag, shot_type) %>%
filter(shot_type == '2PT Field Goal') %>%
summarise(made_2ptshot = length(name)) %>%
group_by(name) %>%
mutate(total_2ptshots = sum(made_2ptshot)) %>%
filter(shot_made_flag == 'made_shot') %>%
select(c(1, 4, 5)) %>%
mutate(perc_2ptmade = made_2ptshot/total_2ptshots)

shots_data1 %>%
group_by(name, shot_made_flag, shot_type) %>%
filter(shot_type == '3PT Field Goal') %>%
summarise(made_3ptshot = length(name)) %>%
group_by(name) %>%
mutate(total_3ptshots = sum(made_3ptshot)) %>%
filter(shot_made_flag == 'made_shot') %>%
select(c(1, 4, 5)) %>%
mutate(perc_3ptmade = made_3ptshot/total_3ptshots)
```

```{r Question 6.1}
shotdistprop <- shots_data1 %>%
group_by(shot_distance, shot_made_flag) %>%
summarise(made_shots = length(shot_distance)) %>%
group_by(shot_distance) %>%
mutate(total_shots = sum(made_shots)) %>%
filter(shot_made_flag == 'made_shot') %>%
select(c(1, 3:4)) %>%
mutate(made_shot_prop = made_shots/total_shots) %>%
select(c(1, 4))
```

```{r Question 6.2}
ggplot(shotdistprop, aes(x = shot_distance, y = made_shot_prop)) + geom_point()
```

Question 6.2 short answers:
I observe that the first three shot distances clearly have a higher rate of effectiveness. The 51 foot shot has a 100% success rate, because it was only shot once and it happened to go in, so it's an outlier. Beyond a shot distance of 3, the made shot percentage seems to be going up until around the 20 shot distance range. This may be because players practice farther shots as those opportunities are easier to find  in a game. I can't confirm that the shorter the distance, the more effective the shots. Since an NBA basketball court is 94 feet, a threshhold beyond which the chance of making a successful shot is anything beyond 94 feet. Any shot within the bounds of the court has a possibility of success. The distances that tend to have a successful shot percentage of 50% or more are distances: 1, 2, 3, 13, 16, 21, 30, 51. 

```{r Question 7}
shotsbymin <- shots_data1 %>%
group_by(name, minute) %>%
summarise(total_shots = length(name))

ggplot(shotsbymin, aes(x = minute, y = total_shots)) + geom_point(col = 'skyblue1') + geom_path(col = 'skyblue1') + scale_x_continuous(limits = c(0, 48), labels = c(1, 12, 24, 36, 48), breaks = c(1, 12, 24, 36, 48)) + geom_rect(xmin=0, xmax=12, ymin=0, ymax=60, alpha = 0.005) + theme_minimal() + geom_rect(xmin = 24, xmax = 36, ymin = 0, ymax = 60, alpha = 0.005) + facet_wrap(~ name) 
```